---
title: "Statistical Inference Theory - Lab Report"
output: 
  html_notebook: default
  pdf_document: default
---

# Binomial Distribution

The binomial distribution is a discrete probability distribution. It describes the outcome of n independent trials in an experiment. Each trial is assumed to have only two outcomes, either success or failure. If the probability of a successful trial is p, then the probability of having x successful outcomes in an experiment of n independent trials is as follows.

f(x) = nCx * p^x * (1 - p)^(n-x)

## Problem

Suppose there are twelve multiple choice questions in an English class quiz. Each question has five possible answers, and only one of them is correct. Find the probability of having four or less correct answers if a student attempts to answer every question at random.

### Solution

```{r}
pbinom(4, size = 12, prob = 0.2)
```
# Poisson Distribution

The Poisson distribution is the probability distribution of independent event occurrences in an interval. If λ is the mean occurrence per interval, then the probability of having x occurrences within a given interval is:  

f(x) = λ^x * e^(- λ) / x!

## Problem

If there are twelve cars crossing a bridge per minute on average, find the probability of having seventeen or more cars crossing the bridge in a particular minute.

### Solution

```{r}
ppois(16, lambda = 12, lower = FALSE)
```
# Normal Distribution

The normal distribution is defined by the following probability density function, where μ is the population mean and σ^2 is the variance.

f(x) = e^( -(x - μ)^2 / 2 * σ^2) / σ * sqrt(2 * pi)

## Problem

Assume that the test scores of a college entrance exam fits a normal distribution. Furthermore, the mean test score is 72, and the standard deviation is 15.2. What is the percentage of students scoring 84 or more in the exam?

### Solution

```{r}
pnorm(84, mean = 72, sd = 15.2, lower.tail = FALSE)
```
```{r}

```
```{r}

```
```{r}

```
```{r}

```
```{r}

```
```{r}

```
```{r}

```
```{r}

```
```{r}

```
