---
title: "Statistical Inference Theory - Lab 9"
output: html_notebook
---

#### CB.SC.I5DAS20032

## Problem 1

In order to determine the possible effect of a chemical treatment on the rate of germination of cotton seeds, a pot culture was conducted. The results are given below. Significance at 0.05.

## Solution

```{r}
Sample2 = matrix(c(118, 120, 22, 40), nrow = 2, ncol = 2)

Sample2
```

```{r}
chisq.test(Sample2)
```
As p = 0.06596 > 0.05, h0: Attributes are independent is **not rejected**.


## Problem 2

The severity of a disease and blood group were studied in a research project.

```{r}
DisxBlood = matrix(c(51, 105, 384, 40, 103, 527, 10, 25, 125, 9, 17, 104), nrow = 3, ncol = 4)

DisxBlood
```
### Solution

```{r}
chisq.test(DisxBlood)
```
As p = 0.2003 > 0.05, h0: Attributes are independent is **not rejected**.

```{r}
qchisq(0.95, df = 6)
```
As chisq = 12.237 < chisq(table) = 12.59159, h0: Attributes are independent is **not rejected**.

## Problem 3

A public opinion poll surveyed a simple random sample of 1000 voters.

```{r}
GenxParty = matrix(c(220, 270, 170, 320, 70, 70), nrow = 2, ncol = 3)

GenxParty
```
### Solution

```{r}
chisq.test(GenxParty)
```
As p = 0.0003688 < 0.05, h0: Attributes are independent is **rejected**.

```{r}
qchisq(0.95, df = 2)
```
As chisq = 15.81 < chisq(table) = 5.991465, h0: Attributes are independent is **rejected**.
